---
title: "Proposal: The Effect of U.S. Military Aid on FDI"
author: "Ben Horvath"
date: "November 1, 2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    theme: null
    css: ../../static/architect.css
    template: ../../static/architect.html
    pandoc_args: [
      "--mathjax", "",
      "--variable", "mathjax-url:https://mathjax.rstudio.com/latest/MathJax.js?config=TeX-AMS-MML_HTMLorMML"]
fontsize: 12pt
fontfamily: mathpazo
fontfamilyoptions: osf,sc
linestretch: 1.1
header-includes:
  \usepackage{eulervm}
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# setwd("~/Dropbox/Documents/school/grad/606/fdi_arms_exports")
```

Load libraries:

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(Hmisc)
library(knitr)
library(readxl)
library(stringr)
library(tidyr)
```

# Introduction

I took a 200-level 'methods in political science' class in the spring semester of 2009, as part of an undergraduate degree in political science. One assignment was to write a proposal for empirical research, though we weren't required to necessarily carry out the analysis.

Biglaiser and DeRouen (2007) and Little and Leblang (2004) had found that 'the presence of U.S. troops serves as a "catalyst" for U.S. outgoing foreign direct investment (FDI), that is, FDI follows the flag.' My proposal, 'The Effect of U.S. Military Aid on Foreign Direct Investment Decisions,' was to test if U.S. arms exports had a similar effect on FDI as U.S. troops. I located the appropriate data, and noted a number of other variables that would need to be controlled for: alliances, the presence of conflict, the Cold War and outliers like Vietnam, regime type (democratic, autocratic, etc.), and population size.

For methodology, I wrote, simply, 'To test these associations, some kind of regression would be used, with an appropriate significance test.' I don't believe I really knew what a regression was, I'd just noticed how popular it was in the political science literature.

Nine years later, I am much more sophisticated statistically, and would like to see how my college sophomore intuition faired.

#### References 

* Glenn Biglaiser and Karl DeRouen, Jr. (2007), 'Following the flag: Troop deployment and US foreign direct investment,' _International Studies Quarterly_ 51, no. 4: 835-854.

* Andrea Little and David Leblang (2004), 'Military securities: Financial flows and the deployment of US troops,' in _Annual Meeting of the American Political Science Association_, pp. 2-5.


# Research Questions

What is the effect of an increase in U.S. arms exports to a country's incoming U.S. foreign direct investment?



# Cases

Each row will be attributes associated with a single year for a single country: `(year, country)`.

_Note:_ I realize the basic regression we're going to perform is not ideal for this kind of cross-sectional longitudinal data set. It clearly violates, at least, the assumption of independence between observations. However, I'd like to see why it doesn't work for myself, on a concrete dataset I understand. I'd also like to compare this basic regression's performance against the 'proper' way as well as the methodology of the studies referenced above, as a bonus.



# Response and Explanatory Variables

The response variable is incoming U.S. FDI from a country, measured in USD.

The explanatory variable we are most interested in is U.S. military aid (probably lagged a year).

The studies referenced above include a few control variables, including population, existence of a conflict in that year and country, type of regime (democracy, dictatorship, etc.), alliance statuses, distance between countries, and GDP. These variables will also be lagged a year for modeling.

All of this data is easily assembled, if you know where to look, so I would like to include those as well.



# Data Sources

* __FDI__. The OECD provides FDI data for U.S. outflows on its website, from 2003 to 2013. These years will have to bound this study temporally:  https://stats.oecd.org/index.aspx?DataSetCode=FDI_FLOW_PARTNER

* __Arms transfers__. The Stockholm International Peace Research Institute maintains a database of arms transfers: https://www.sipri.org/databases/armstransfers. The value has been 'normalized' by the researchers themselves to account for fluctuations in the market value of weapons as well as allowing comparability between, e.g., 100 assault rifles and 2 large artilleries.

* __Yearly Population__. This is available for most countries on a yearly basis via the UN: https://population.un.org/wpp/Download/Standard/Population/

* __Presence of Conflict__. Political scientists testing hypotheses on armed conflict frequently make use of the Armed Conflict dataset, available at: http://ucdp.uu.se/downloads/#d3. This dataset contains a lot of data, but I am just going to use a dichotomous variable: 0 for no conflict, 1 for conflict.

* __Regime Type__. This is available in one of the most popular political science datasets, the Polity dataset. It encodes regime type in a range from perfectly democratic to perfectly autocratic for most countries from 1800 on. Specifically I will use the Polity2 variable: http://www.systemicpeace.org/inscrdata.html. See also the user manual: http://www.systemicpeace.org/inscr/p4manualv2017.pdf.

* __Alliances__. The Correlates of War project maintains another popular data set encoding international alliances in 'dyadic' form a year-to-year basis: http://www.correlatesofwar.org/data-sets/formal-alliances.

* __Distance__. Kristian Gleditsch developed a data set containing the distance between capital cities, which we'll use to proxy distance: http://ksgleditsch.com/data-5.html, using this system of country codes: http://ksgleditsch.com/statelist.html

* __GDP__. Where else but the World Bank?:   https://data.worldbank.org/indicator/NY.GDP.MKTP.CD



# Data Collection and Preparation

## Collecting Data

The goal will be to combine this data in a clean format, for as many countries as possible between 2003 and 2013. Filter the dataset to only include outflow numbers.

### FDI

```{r}
fdi <- read.csv('../data/raw/FDI_FLOW_PARTNER_28102018214703504.csv',
                stringsAsFactors=FALSE)
colnames(fdi) <- tolower(colnames(fdi))

# we are only interested in outflow, from the U.S. to other countries
fdi <- fdi %>% 
    filter(reporting.country == 'United States',
           type.of.fdi == 'Outward',
           cur == 'USD')
```

The data also includes various aggregated rows, including regions like GULF ARABIAN COUNTRIES, these must be filtered out as well.

```{r}
`%not in%` <- function (x, table) is.na(match(x, table, nomatch=NA_integer_))

aggregations <- c('ACP countries', 'AFRICA', 'African ACP countries', 'AMERICA', 'ASEAN countries', 'ASIA', 'BALTIC COUNTRIES', 'BLEU', 'Caribbean ACP countries', 'CENTRAL AMERICA', 'EU15', 'EU25', 'EUROPE', 'Other territories', 'GULF ARABIAN COUNTRIES', 'MERCOSUR', 'Mediterranean Basin countries', 'NAFTA', 'NEAR AND MIDDLE EAST', 'NICs1', 'NICs2A', 'NICs2LA', 'NORTH AFRICA', 'NORTH AMERICA', 'OCEANIA & POLAR REGIONS', 'OECD', 'OPEC countries', 'OTHER AFRICAN COUNTRIES (Excluding North Africa)', 'OTHER ASIAN COUNTRIES (Excluding Near and Middle East)', 'Other Central/Eastern Europe', 'NEAR AND MIDDLE EAST COUNTRIES OTHER THAN GULF STATES', 'Pacific ACP countries', 'SOUTH AMERICA', 'TOTAL WORLD', 'EUROPE Unallocated', 'EUROPE (Excluding OECD countries)', 'AFRICA Unallocated', 'North Africa - Unallocated', 'Africa excluding North African countries- Unallocated', 'AMERICA Unallocated', 'Central America - Unallocated', 'CENTRAL AMERICA (Excluding OECD countries)', 'South America - Unallocated', 'ASIA Unallocated', 'ASIA (Excluding OECD countries)', 'Near and Middle East - Unallocated', 'Gulf Arabian countries - Unallocated', 'Near and Middle East countries excluding Gulf states - Unallocated', 'Other Asian countries - Unallocated', 'OTHER ASIAN COUNTRIES (Excluding Near and Middle East and OECD countries)', 'Oceania & Polar Regions - Unallocated', 'OCEANIA & POLAR REGIONS (Excluding OECD countries)', 'TOTAL WORLD Unallocated', 'TOTAL WORLD (Excluding OECD countries)', 'OECD - Unallocated', 'NORTH AMERICA (Excluding OECD countries)', 'AMERICA (Excluding OECD countries)', 'Economic zones', 'EU27', 'SOUTH AMERICA (Excluding OECD countries)', 'NEAR AND MIDDLE EAST (Excluding OECD countries)', 'CIS countries', 'EFTA')

fdi <- fdi %>% filter(partner.country %not in% aggregations)
```

Clean up the columns a bit:

```{r}
us_fdi <- fdi %>%
    select(year, partner.country, value) %>%
    mutate(value = value * 1000000) %>%
    arrange(year, partner.country)
colnames(us_fdi) <- c('year', 'country', 'fdi')
rm(fdi)

write.table(us_fdi, '../data/clean/fdi.tsv', row.names=FALSE, sep='\t')

head(us_fdi)
```



### Arms Transfers

```{r}
# not really an Excel file!
arms_raw <- read.csv('../data/raw/TIV-Export-USA-2003-2013.csv.xls', skip=10, 
                 header=TRUE)
arms_raw$Total <- NULL
colnames(arms_raw)[1] <- 'country'

arms_exports <- arms_raw %>%
    gather(year, arms_exports, X2003:X2013, na.rm=TRUE) %>%
    mutate(year=as.integer(str_remove(year, 'X'))) %>%
    arrange(country, year)
colnames(arms_exports)[3] <- 'arms_exports'
rm(arms_raw)

write.table(arms_exports, '../data/clean/arms_exports.tsv', row.names=FALSE, sep='\t')

head(arms_exports)
```



### Population

```{r}
pop_raw <- read_xlsx('../data/raw/WPP2017_POP_F01_1_TOTAL_POPULATION_BOTH_SEXES.xlsx',
                     skip=16, col_names=TRUE) %>%
    select(3, `2003`:`2013`)
colnames(pop_raw)[1] <- 'country'

pop_aggs <- c('WORLD', 'More developed regions', 'Less developed regions', 'Least developed countries', 'Less developed regions, excluding least developed countries', 'Less developed regions, excluding China', 'High-income countries', 'Middle-income countries', 'Upper-middle-income countries', 'Lower-middle-income countries', 'Low-income countries', 'Sub-Saharan Africa', 'AFRICA', 'Middle Africa', 'Eastern Africa', 'Western Sahara', 'Southern Africa', 'ASIA', 'Eastern Asia', 'Southern Asia', 'South-Eastern Asia', 'EUROPE', 'Eastern Europe', 'Northern Europe', 'Southern Europe', 'Western Europe', 'NORTHERN AMERICA', 'OCEANIA', 'Central Asia', 'LATIN AMERICA AND THE CARIBBEAN', 'South-Central Asia', 'Western Africa', 'Northern Africa')

pop <- pop_raw %>% 
    filter(country %not in% pop_aggs) %>%
    gather(year, population, `2003`:`2013`) %>%
    mutate(population = population * 1000,
           year = as.numeric(year))

rm(pop_raw)
write.table(pop, '../data/clean/population.tsv', row.names=FALSE, sep='\t')

head(pop)
```



### Conflict

```{r}
conflict <- read.csv('../data/raw/ucdp-prio-acd-181.csv', 
                         stringsAsFactors=FALSE) %>%
    select(conflict_id, location, year) %>%
    mutate(conflict = 1) %>%
    arrange(conflict_id, location, year)
colnames(conflict)[2] <- 'country'

write.table(conflict, '../data/clean/conflict.tsv', row.names=FALSE, sep='\t')

head(conflict)
```



### Regime Type

```{r}
regime <- read_xls('../data/raw/p4v2017.xls') %>%
    select(country, year, polity2) %>%
    arrange(country, year)

write.table(regime, '../data/clean/regime.tsv', row.names=FALSE, sep='\t')

head(regime)
```



### Alliances

```{r}
alliances <- read.csv('../data/raw/alliance_v4.1_by_dyad_yearly.csv',
                      stringsAsFactors=FALSE) %>%
    filter(state_name1 == 'United States of America',
           year >= 2003,
           year <= 2013) %>%
    select(state_name2, year) %>%
    mutate(alliance = 1)
colnames(alliances)[1] <- 'country'

write.table(alliances, '../data/clean/alliances.tsv', row.names=FALSE, sep='\t')

head(alliances)
```



### Distance

```{r}
countries <- read.table('../data/raw/iisystem.dat', sep='\t',
                        stringsAsFactors=FALSE)

distance <- read.csv('../data/raw/capdist.csv', stringsAsFactors=FALSE) %>%
    filter(ida == 'USA') %>%
    inner_join(countries, by=c('idb'='V2')) %>%
    select(V3, kmdist)
colnames(distance) <- c('country', 'km_dist')

rm(countries)
write.table(distance, '../data/clean/distance.tsv', row.names=FALSE, sep='\t')

head(distance)
```



### GDP

Extracting two variables, absolute GDP and yearly percentage growth:

```{r}
gdp <- read.csv('../data/raw/API_NY.GDP.MKTP.CD_DS2_en_csv_v2_10203569.csv',
                stringsAsFactors=FALSE, skip=4) %>%
    select(Country.Name, X2002:X2013) %>%
    gather(year, gdp, X2002:X2013) %>%
    mutate(year = as.numeric(str_remove(year, 'X'))) %>%
    arrange(Country.Name, year)
colnames(gdp) <- c('country', 'year', 'gdp')

gdp <- gdp %>% 
    group_by(country) %>% 
    mutate(gdp_l1 = lag(gdp, n=1, default=NA)) %>%
    mutate(gdp_perc_growth = (gdp - gdp_l1) / gdp_l1) %>%
    filter(year >= 2003, year <= 2013) %>%
    select(country, year, gdp, gdp_perc_growth)

write.table(gdp, '../data/clean/gdp.tsv', sep='\t', row.names=FALSE)

head(gdp)
```



## Putting the Data Together

```{r}
df <- us_fdi %>%
    inner_join(pop, by=c('country', 'year')) %>%
    left_join(arms_exports, by=c('country', 'year')) %>%
    left_join(conflict,  by=c('country', 'year')) %>%
    left_join(regime, by=c('country', 'year')) %>%
    left_join(alliances, by=c('country', 'year')) %>%
    left_join(distance, by=c('country')) %>%
    left_join(gdp, by=c('country', 'year')) %>%
    select(-conflict_id) %>%
    arrange(country, year)

# Fill in missing variables to indicate absense of exports, etc.
df <- df %>%
    mutate(arms_exports = replace_na(arms_exports, 0),
           conflict = replace_na(conflict, 0),
           alliance = replace_na(alliance, 0))
           

write.table(df, '../data/clean/master_dataset.tsv', row.names=FALSE, sep='\t')

head(df)
```



# Summary Statistics


These are histograms of the two main variables of interest, incoming FDI and arms imports from the U.S. Both are about the same shape, with many countries that have little or no FDI or arms imports, and a long right tail.

```{r}
par(mfrow=c(1,2))
hist(df$fdi, main='FDI', col='gray')
hist(df$arms_exports, main='Arms Imports', col='gray')
```

It might make sense for modeling purposes to take the logs of these variables, or at least of FDI---though some values are negative:

```{r}
par(mfrow=c(1,2))
hist(log(df$fdi), main='FDI Logged', col='gray')
hist(log(df$arms_exports), main='Arms Imports Logged', col='gray')
```

The tremendous skew in FDI is obvious by comparing the mean and the median, almost 2 billion and 10 million, respectively. Standard deviation is also very high, almost 700 million.

```{r}
describe(df$fdi)
```

It occurrs to me it might make sense to standardize FDI by dividing it by a country's population or GDP---something to experiment with.

Just for fun, let's look at the top 10 recipients of FDI from the U.S. (in millions of USD):

```{r}
df %>% group_by(country) %>%
    summarise(fdi = sum(fdi)) %>%
    mutate(fdi = fdi / 1000000) %>%
    arrange(desc(fdi)) %>%
    select(country, fdi) %>%
    top_n(10) %>%
    kable
```

Most are highly developed nations with large GDPs, half of them in Europe.

Arms exports also has a much larger mean than median, the latter at 0, i.e., over half of these year-country units received no arms from the United States.

```{r}
describe(df$arms_exports)
```

The top 10 recipients of U.S. arms over this time period:

```{r}
df %>% group_by(country) %>%
    summarise(arms_exports = sum(arms_exports )) %>%
    arrange(desc(arms_exports )) %>%
    select(country,arms_exports ) %>%
    top_n(10) %>%
    kable
```

This all looks correct. Israel is the largest recipient, and Egypt receives tons of aid under the Camp David Treaty that President Carter negotiated.

Let's look at a scatterplot to view the direct relationship between FDI and arms exports:

```{r}
ggplot(na.omit(df), aes(x=arms_exports, y=fdi)) + 
    geom_point() +
    geom_smooth() +
    ggtitle('Arms Exports and FDI')
```

Let's try taking the log of both variables:

```{r}
ggplot(na.omit(df), aes(x=log(arms_exports), y=log(fdi))) + 
    geom_point() +
    geom_smooth() +
    ggtitle('log(Arms Exports) and log(FDI)')
```

This looks much better! Let's just try one more, dividing both values by population and then taking the log:

```{r}
ggplot(na.omit(df), aes(x=log(arms_exports/population), y=log(fdi/population))) + 
    geom_point() +
    geom_smooth() +
    ggtitle('Population Standardized log(Arms Exports) and log(FDI)')
```

BEAUTIFUL. This graph actually looks so good I feel like I've cheated somewhere?? Those residuals are going to be so normal.



# Misc. Notes for Ben

## TODO

1. Create some kind of standardized mapping of countries so that more of them get passed through all the joins. 

2. Consider some strategies to fill in missing values. E.g., Polity does not assign a regime for Afghanistan for several years because of the conflict there---what makes sense as a way to handle this reasonably without just dropping the rows?

3. Fill all NAs with zeros where NA indicates absence of phenomena.

4. Attempt three models: regular `lm`, `plm` for panel data, and original two-step least square regression of the original studies, paying particular attention to showing the violation of `lm` assumptions and how the latter two correct this.

## Modeling

Note on the original studies' two-stage least squares regression:

1. Authors first developed a _troop_ model: 

$$troops \sim conflict_{-1} + alliance_{-1} + polity_{-1} + warsaw\_pact_{-1} + cold\_war_{-1} +$$
$$log(pop_{-1}) + reagan_{-1} + south\_korea + vietnam + philippines$$

2. Then plugged the results of that model into a _trade_ model with some other variables:

$$trade \sim troops + growth_{-1} + gdp_{-1} + distance + alliance_{-1}$$

where -1 subscript indicates lag of 1 year; both models including intercepts.

They do this because they suspect trade and presence of U.S. troops are endogenous to eachother; in stastical langauge there will be a correlation with model errors.


